---
title: "RPE FUN ANOVA"
author: "Ian Brown"
date: "2025-02-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r chunk2, echo=FALSE, include=FALSE}
library(tidyverse)
library(ggpubr)
library(rstatix)
```

```{r chunk3, echo=FALSE}
RPE.FUN.wide <- read.csv('Data/RPE_FUN_AVG.csv', header = T)
RPE.FUN.wide <- na.omit(RPE.FUN.wide)
RPE.FUN.wide <- RPE.FUN.wide[RPE.FUN.wide$WEEK < 11, ]
attach(RPE.FUN.wide)

RPE.FUN <- RPE.FUN.wide %>%
  gather(key = "outcome", value = "score",
         HIIT.RPE, HIIT.FUN, PLAY.RPE, PLAY.FUN) %>%
  mutate(outcome = as.factor(outcome)) %>%
  mutate(SUBJECT = as.factor(SUBJECT)) %>%
  mutate(WEEK = as.factor(WEEK))
RPE.FUN <- RPE.FUN[, -c(4,5,6,8:24)]
attach(RPE.FUN)
```

### One-way ANOVA

###### Summary Stats

```{r chunk4, echo=FALSE}
RPE.FUN %>%
  group_by(outcome) %>%
  get_summary_stats(score, type = "mean_sd")

bxp <- ggboxplot(RPE.FUN, x = "outcome", y = "score", add = "point")
bxp
```

###### Assumptions

```{r chunk5, echo=FALSE}
RPE.FUN %>%
  group_by(outcome) %>% 
  identify_outliers(score) # no extreme outliers
RPE.FUN %>%
  group_by(outcome) %>%
  shapiro_test(score) # does not pass shapiro test but n > 50. Check QQ-plot
ggqqplot(RPE.FUN, "score", title = "Non-Transformed", facet.by = "outcome")
```

There are no extreme outliers. Did not pass Shapiro-Wilk test, however, due to n>50 this test becomes very sensitive to minor deviations, so QQ-plots were created. Normality assumption based on QQ-plots is on the verge of not being met. Did various transformations of "score" to see if I could get a better distribution. Sphericity is check during computation. All transformations were either about the same or much worse. Analysis run with non-transformed data.

###### Computation

```{r chunk6, echo=FALSE}
anova.score <- anova_test(data = RPE.FUN, dv = score,
                          wid = c(SUBJECT,WEEK), within = outcome)
data.frame(get_anova_table(anova.score))
```

###### Post-hoc tests

```{r chunk7, echo=FALSE}
pwc <- RPE.FUN %>%
  pairwise_t_test(
    score ~ outcome, paired = TRUE,
    p.adjust.method = "bonferroni"
  )
data.frame(pwc)
```

###### Box Plots with p-values

```{r chunk8, echo=FALSE}
pwc <- pwc %>% add_xy_position(x = "outcome")
bxp +
  stat_pvalue_manual(pwc) +
  labs(subtitle = get_test_label(anova.score, detailed = T),
       caption = get_pwc_label(pwc))
```

### Two-way ANOVA between WEEK (time) and outcome

###### Summary Stats

```{r chunk9, echo=FALSE}
RPE.FUN %>%
  group_by(outcome, WEEK) %>%
  get_summary_stats(score, type = "mean_sd")

bxp <- ggboxplot(RPE.FUN,
                 x = "WEEK", y = "score", add = "point",
                 color = "outcome")
bxp
```

###### Assumptions

```{r chunk10, echo=FALSE}
RPE.FUN %>%
  group_by(outcome, WEEK) %>% 
  identify_outliers(score) # 
RPE.FUN %>%
  group_by(outcome, WEEK) %>%
  shapiro_test(score) # does not pass shapiro test but n > 50. Check QQ-plot
ggqqplot(RPE.FUN, "score", title = "Non-Transformed")+
  facet_grid(outcome ~ WEEK, labeller = "label_both")
```

There are 2 extreme outliers (SUBJECT 56 WEEK 5 and SUBJECT 1 WEEK 10). Similar to one-way ANOVA, most groups do not pass the Shapiro-Wilk test for normality. However, there are some groups that do pass it. Most QQ-plots looks similar to those groups that passed the Shapiro-Wilk test.

###### Computation

```{r chunk11, echo=FALSE}
anova.score <- anova_test(data = RPE.FUN, dv = score,
                          wid = SUBJECT, within = c(WEEK, outcome))
data.frame(get_anova_table(anova.score))
```

###### Post-hoc tests

```{r chunk12, echo=FALSE}
pwc <- RPE.FUN %>%
  group_by(WEEK) %>%
  pairwise_t_test(score ~ outcome, paired = TRUE,
                  p.adjust.method = "bonferroni")
data.frame(pwc)
```

###### Box-plots with p-values

```{r chunk13, echo=FALSE}
pwc <- pwc %>% add_xy_position(x = "WEEK")
bxp +
  stat_pvalue_manual(pwc, tip.length = 0, hide.ns = TRUE) +
  labs(subtitle = get_test_label(anova.score, detailed = T),
       caption = get_pwc_label(pwc))
```


