---
title: "RPE-FUN ANALYSIS"
author: "Ian Brown"
date: "2025-02-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r chunk2, echo=FALSE, include=FALSE}
library(tidyr)
library(dplyr)
library(rstatix)
library(nlme)
library(leaps)
```

### Data Preparation and Analysis Background

Data were averaged between both sessions of HIIT and PLAY. The time spent in the 5 HR zones were split between Vigorous, Moderate, and Light (5/4, 3/2, and 1 respectively). All subjects with missing values were excluded at this time.

```{r chunk3, echo=FALSE, include=FALSE}
RPE.FUN <- read.csv('Data/RPE_FUN_AVG.csv', header = T) # 897 obs.
RPE.FUN <- na.omit(RPE.FUN)
# RPE.FUN <- RPE.FUN %>% drop_na(HIIT.RPE | HIIT.FUN | PLAY.RPE | PLAY.FUN) # 383 obs.

attach(RPE.FUN)

# Change class types as needed
RPE.FUN <- RPE.FUN %>%
  mutate(SUBJECT = as.character(SUBJECT))
RPE.FUN <- RPE.FUN %>%
  mutate(SEX = as.factor(SEX))
RPE.FUN <- RPE.FUN %>%
  mutate(WEEK = as.character(WEEK))
RPE.FUN <- RPE.FUN %>%
  mutate(LESSON.PLAN = as.factor(LESSON.PLAN))
RPE.FUN <- RPE.FUN %>%
  mutate(HIIT.PLAY.1 = as.factor(HIIT.PLAY.1))

attach(RPE.FUN)

# Fill remaining NAs with column medians
# RPE.FUN <- RPE.FUN %>% mutate(across(c(BASE.HR,
                                       # HIIT.SEC, HIIT.VIG.SEC, HIIT.MOD.SEC, HIIT.LIGHT.SEC,
                                       # HIIT.VIG.PERC, HIIT.MOD.PERC, HIIT.LIGHT.PERC,
                                       # PLAY.SEC, PLAY.VIG.SEC, PLAY.MOD.SEC, PLAY.LIGHT.SEC,
                                       # PLAY.VIG.PERC, PLAY.MOD.PERC, PLAY.LIGHT.PERC),
                                     # ~replace_na(., median(., na.rm = TRUE))))

# attach(RPE.FUN)
```

### 2-Sided t-tests with Bonferroni adjusted p-values

```{r chunk4, echo=FALSE}
t1 <- t.test(HIIT.SEC, PLAY.SEC, paired = TRUE)
t2 <- t.test(HIIT.RPE, PLAY.RPE, paired = TRUE)
t3 <- t.test(HIIT.FUN, PLAY.FUN, paired = TRUE)
t4 <- t.test(HIIT.VIG.PERC, PLAY.VIG.PERC, paired = TRUE)
t5 <- t.test(HIIT.MOD.PERC, PLAY.MOD.PERC, paired = TRUE)

M.subset <- subset(RPE.FUN, SEX == "M")
F.subset.full <- subset(RPE.FUN, SEX == "F")
F.subset <- sample_n(F.subset.full, 109)

t6 <- t.test(M.subset$HIIT.RPE, F.subset$HIIT.RPE, paired = T)
t7 <- t.test(M.subset$HIIT.FUN, F.subset$HIIT.FUN, paired = T)
t8 <- t.test(M.subset$PLAY.RPE, F.subset$PLAY.RPE, paired = T)
t9 <- t.test(M.subset$PLAY.FUN, F.subset$PLAY.FUN, paired = T)

# Vector of all p-values
d1 <- data.frame(test = "Time HIIT vs PLAY", est = t1$estimate, raw_p=t1$p.value)
d2 <- data.frame(test = "RPE HIIT vs PLAY", est = t2$estimate, raw_p=t2$p.value)
d3 <- data.frame(test = "FUN HIIT vs PLAY", est = t3$estimate, raw_p=t3$p.value)
d4 <- data.frame(test = "Vig % HIIT vs PLAY", est = t4$estimate, raw_p=t4$p.value)
d5 <- data.frame(test = "Mod % HIIT vs PLAY", est = t5$estimate, raw_p=t5$p.value)
d6 <- data.frame(test = "RPE HIIT Sex", est = t6$estimate, raw_p=t6$p.value)
d7 <- data.frame(test = "FUN HIIT Sex", est = t7$estimate, raw_p=t7$p.value)
d8 <- data.frame(test = "RPE PLAY Sex", est = t8$estimate, raw_p=t8$p.value)
d9 <- data.frame(test = "FUN PLAY Sex", est = t9$estimate, raw_p=t9$p.value)

res <- rbind(d1,d2,d3,d4,d5,d6,d7,d8,d9)
res$bon <- p.adjust(res$raw_p, method = "bonferroni")
res
```

### Linear Mixed Effect Models

####### HIIT.RPE

```{r chunk5, echo=FALSE}
m1 <- lme(HIIT.RPE ~ PLAY.FUN, data = RPE.FUN,
          random = ~ 1|SUBJECT,
          control = lmeControl(opt = 'optim'))
m2 <- lme(HIIT.RPE ~ PLAY.RPE, data = RPE.FUN,
          random = ~ 1|SUBJECT,
          control = lmeControl(opt = 'optim'))
m3 <- lme(HIIT.RPE ~ HIIT.FUN, data = RPE.FUN,
          random = ~ 1|SUBJECT,
          control = lmeControl(opt = 'optim'))
m4 <- lme(HIIT.RPE ~ HIIT.MOD.PERC+HIIT.VIG.PERC, data = RPE.FUN,
          random = ~ 1|SUBJECT,
          control = lmeControl(opt = 'optim'))
m5 <- lme(HIIT.RPE ~ AGE, data = RPE.FUN,
          random = ~ 1|SUBJECT,
          control = lmeControl(opt = 'optim'))

tut1 <- summary(m1)
tabl1 <- tut1$tTable
tut2 <- summary(m2)
tabl2 <- tut2$tTable
tut3 <- summary(m3)
tabl3 <- tut3$tTable
tut4 <- summary(m4)
tabl4 <- tut4$tTable
tut5 <- summary(m5)
tabl5 <- tut5$tTable

tabl1
tabl2
tabl3
tabl4
tabl5
```



```{r chunk6, echo=FALSE}
reg.fit <- regsubsets(HIIT.RPE ~ HIIT.FUN+PLAY.RPE+PLAY.FUN+
                      BASE.HR+HIIT.VIG.PERC+HIIT.MOD.PERC+PLAY.VIG.PERC+PLAY.MOD.PERC+
                      AGE+SEX+HEIGHT..cm.+WEIGHT..kg.+LESSON.PLAN,
                      nvmax = 30,
                      data = RPE.FUN)
reg.summary <- summary(reg.fit)

w_cp<-which.min(reg.summary$cp)
plot(reg.summary$cp, xlab = "Number of Variables",
    ylab = "Cp", type = "l",main="Pick based on Cp")
points(w_cp, reg.summary$cp[w_cp], col = "red", cex = 2,
    pch = 20)

n <- length(RPE.FUN$HIIT.RPE)
p <- apply(reg.summary$which, 1, sum)
aic <- reg.summary$bic - log(n) * p + 2 * p
w_a<-unname(which.min(aic))
plot(aic, xlab = "Number of Variables",
    ylab = "AIC", type = "l",main="Pick based on AIC")
points(w_a, aic[w_a], col = "blue", cex = 2,
    pch = 20)

w_b<-which.min(reg.summary$bic)
plot(reg.summary$bic, xlab = "Number of Variables",
    ylab = "BIC", type = "l",main="Pick based on BIC")
points(w_b, reg.summary$bic[w_b], col = "green", cex = 2,
    pch = 20)

w_r<-which.max(reg.summary$adjr2)
plot(reg.summary$adjr2, xlab = "Number of Variables",
    ylab = "adj r2", type = "l",main="Pick based on adjusted Rsquared")
points(w_r, reg.summary$adjr2[w_r], col = "orange", cex = 2,
    pch = 20)

w_r2<-which.max(reg.summary$rsq)
plot(reg.summary$rsq, xlab = "Number of Variables",
    ylab = "r2", type = "l",main="Pick based on Rsquared")
points(w_r2, reg.summary$rsq[w_r2], col = "purple", cex = 2,
    pch = 20)

w_cp
w_a
w_b
w_r
w_r2
```



```{r chunk7, echo=FALSE}
m <- lme(HIIT.RPE ~ 
          PLAY.RPE+
          # PLAY.FUN+
          HIIT.FUN+
          # HIIT.VIG.PERC+HIIT.MOD.PERC+PLAY.VIG.PERC+HIIT.MOD.PERC+
          # SEX+WEIGHT..kg.+
          LESSON.PLAN,
          data = RPE.FUN,
          random = ~ 1|SUBJECT,
          control = lmeControl(opt = 'optim'))
tut <- summary(m)
tabl <- tut$tTable
tabl
```

####### HIIT.FUN

```{r chunk8, echo=FALSE}
reg.fit <- regsubsets(HIIT.FUN ~ HIIT.RPE+PLAY.RPE+PLAY.FUN+
                      BASE.HR+HIIT.VIG.PERC+HIIT.MOD.PERC+PLAY.VIG.PERC+PLAY.MOD.PERC+
                      AGE+SEX+HEIGHT..cm.+WEIGHT..kg.+LESSON.PLAN,
                      nvmax = 30,
                      data = RPE.FUN)
reg.summary <- summary(reg.fit)
reg.summary
```

```{r}
w_cp<-which.min(reg.summary$cp)
plot(reg.summary$cp, xlab = "Number of Variables",
    ylab = "Cp", type = "l",main="Pick based on Cp")
points(w_cp, reg.summary$cp[w_cp], col = "red", cex = 2,
    pch = 20)

n <- length(RPE.FUN$HIIT.RPE)
p <- apply(reg.summary$which, 1, sum)
aic <- reg.summary$bic - log(n) * p + 2 * p
w_a<-unname(which.min(aic))
plot(aic, xlab = "Number of Variables",
    ylab = "AIC", type = "l",main="Pick based on AIC")
points(w_a, aic[w_a], col = "blue", cex = 2,
    pch = 20)

w_b<-which.min(reg.summary$bic)
plot(reg.summary$bic, xlab = "Number of Variables",
    ylab = "BIC", type = "l",main="Pick based on BIC")
points(w_b, reg.summary$bic[w_b], col = "green", cex = 2,
    pch = 20)

w_r<-which.max(reg.summary$adjr2)
plot(reg.summary$adjr2, xlab = "Number of Variables",
    ylab = "adj r2", type = "l",main="Pick based on adjusted Rsquared")
points(w_r, reg.summary$adjr2[w_r], col = "orange", cex = 2,
    pch = 20)

w_r2<-which.max(reg.summary$rsq)
plot(reg.summary$rsq, xlab = "Number of Variables",
    ylab = "r2", type = "l",main="Pick based on Rsquared")
points(w_r2, reg.summary$rsq[w_r2], col = "purple", cex = 2,
    pch = 20)

w_cp
w_a
w_b
w_r
w_r2
```



```{r chunk9}
m <- lme(HIIT.FUN ~ 
          PLAY.FUN+
          HIIT.RPE+
          # HIIT.VIG.PERC+HIIT.MOD.PERC+PLAY.VIG.PERC+HIIT.MOD.PERC+
          # SEX+WEIGHT..kg.+
          LESSON.PLAN,
          data = RPE.FUN,
          random = ~ 1|SUBJECT,
          control = lmeControl(opt = 'optim'))
tut <- summary(m)
tabl <- tut$tTable
tabl
```

####### PLAY.RPE

```{r chunk10}
reg.fit <- regsubsets(PLAY.RPE ~ HIIT.FUN+HIIT.RPE+PLAY.FUN+
                      BASE.HR+HIIT.VIG.PERC+HIIT.MOD.PERC+PLAY.VIG.PERC+PLAY.MOD.PERC+
                      AGE+SEX+HEIGHT..cm.+WEIGHT..kg.+LESSON.PLAN,
                      nvmax = 30,
                      data = RPE.FUN)
reg.summary <- summary(reg.fit)
reg.summary
```

w_cp<-which.min(reg.summary$cp)
plot(reg.summary$cp, xlab = "Number of Variables",
    ylab = "Cp", type = "l",main="Pick based on Cp")
points(w_cp, reg.summary$cp[w_cp], col = "red", cex = 2,
    pch = 20)

n <- length(RPE.FUN$HIIT.RPE)
p <- apply(reg.summary$which, 1, sum)
aic <- reg.summary$bic - log(n) * p + 2 * p
w_a<-unname(which.min(aic))
plot(aic, xlab = "Number of Variables",
    ylab = "AIC", type = "l",main="Pick based on AIC")
points(w_a, aic[w_a], col = "blue", cex = 2,
    pch = 20)

w_b<-which.min(reg.summary$bic)
plot(reg.summary$bic, xlab = "Number of Variables",
    ylab = "BIC", type = "l",main="Pick based on BIC")
points(w_b, reg.summary$bic[w_b], col = "green", cex = 2,
    pch = 20)

w_r<-which.max(reg.summary$adjr2)
plot(reg.summary$adjr2, xlab = "Number of Variables",
    ylab = "adj r2", type = "l",main="Pick based on adjusted Rsquared")
points(w_r, reg.summary$adjr2[w_r], col = "orange", cex = 2,
    pch = 20)

w_r2<-which.max(reg.summary$rsq)
plot(reg.summary$rsq, xlab = "Number of Variables",
    ylab = "r2", type = "l",main="Pick based on Rsquared")
points(w_r2, reg.summary$rsq[w_r2], col = "purple", cex = 2,
    pch = 20)

w_cp
w_a
w_b
w_r
w_r2
```



```{r chunk11}

```

####### PLAY.FUN

```{r chunk12}
reg.fit <- regsubsets(PLAY.FUN ~ HIIT.FUN+PLAY.RPE+HIIT.RPE+
                      BASE.HR+HIIT.VIG.PERC+HIIT.MOD.PERC+PLAY.VIG.PERC+PLAY.MOD.PERC+
                      AGE+SEX+HEIGHT..cm.+WEIGHT..kg.+LESSON.PLAN,
                      data = RPE.FUN)
reg.summary <- summary(reg.fit)

w_cp<-which.min(reg.summary$cp)
plot(reg.summary$cp, xlab = "Number of Variables",
    ylab = "Cp", type = "l",main="Pick based on Cp")
points(w_cp, reg.summary$cp[w_cp], col = "red", cex = 2,
    pch = 20)

n <- length(RPE.FUN$HIIT.RPE)
p <- apply(reg.summary$which, 1, sum)
aic <- reg.summary$bic - log(n) * p + 2 * p
w_a<-unname(which.min(aic))
plot(aic, xlab = "Number of Variables",
    ylab = "AIC", type = "l",main="Pick based on AIC")
points(w_a, aic[w_a], col = "blue", cex = 2,
    pch = 20)

w_b<-which.min(reg.summary$bic)
plot(reg.summary$bic, xlab = "Number of Variables",
    ylab = "BIC", type = "l",main="Pick based on BIC")
points(w_b, reg.summary$bic[w_b], col = "green", cex = 2,
    pch = 20)

w_r<-which.max(reg.summary$adjr2)
plot(reg.summary$adjr2, xlab = "Number of Variables",
    ylab = "adj r2", type = "l",main="Pick based on adjusted Rsquared")
points(w_r, reg.summary$adjr2[w_r], col = "orange", cex = 2,
    pch = 20)

w_r2<-which.max(reg.summary$rsq)
plot(reg.summary$rsq, xlab = "Number of Variables",
    ylab = "r2", type = "l",main="Pick based on Rsquared")
points(w_r2, reg.summary$rsq[w_r2], col = "purple", cex = 2,
    pch = 20)

w_cp
w_a
w_b
w_r
w_r2
```



```{r chunk13}

```



```{r chunk14}

```